---
title: "p8015_hw2_cmh2275"
author: "cmh2275"
date: "2026-10-03"
output: github_document
---

# Problem 1 

## import data

```{r}
library(tidyverse)


options(tibble.print_min = 5)

transit_df = read_csv (file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
                       col_types = cols(Route8 ="c", Route9 = "c", Route10 = "c", Route11 = "c"), 
                       na=c("NA",".","")) |>
        janitor::clean_names() |>
  select(
    line, station_name, station_latitude, station_longitude, 
    starts_with("route"), entry, exit_only, vending, entrance_type, 
    ada) |> 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
  
```

## Paragraph re: data

This dataset includes information about the NYC transit station's entrances and exits across locations and routes. In additon to informaiton about the division, line, station names, station locations, and routes, it provides information about entrance types, entry, exit, vending, staffing and staff hours, ada compliance, if there is free crossover, cross streets, and entrance locations, with a total of 32 variables. The dataset was reduced to include 19 variables: line, station name, station latitude, station longitude, routes 1-11, vending, entrance type, entry, and ada compliance.The entry variable was converted from a character yes/no variable to a logical variable. There are 19 columns and 1868 rows 

## questions

```{r}
transit_df |> 
  select(station_name, line) |> 
  distinct()
```

*There are 465 distinct stations.

```{r}
transit_df |> 
  filter(ada == TRUE) |> 
  select(station_name, line) |> 
  distinct()
```

* 85 Stations are ADA compliant. 

```{r}
transit_df|> 
  filter(vending == "NO") |> 
  pull(entry) |> 
  mean()
```
* The proportion of stations entrances/exits without vending that allow entrance is 0.377. 
```{r}
transit_df |> 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") |> 
  filter(route == "A") |> 
  select(station_name, line) |> 
  distinct()

transit_df |> 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") |> 
  filter(route == "A", ada == TRUE) |> 
  select(station_name, line) |> 
  distinct()
```
* Sixty distinct stations serve the A train. 
* Seventeen of the sixty stations that serve the A train are ADA compliant.  

# Problem 2 
## read in data 
```{r}
library(tidyverse)
library(readxl)

mrtrash_df = (read_excel("./data/202610 Trash Wheel Collection Data.xlsx", sheet = 1, range = "A2:N742", col_names = TRUE ))|>
  janitor::clean_names()|>
  mutate(vessel = "mr", sports_balls = as.integer(sports_balls), year = as.double(year))|>
  relocate(vessel)


```
## read additional data
```{r}
proftrash_df = (read_excel("./data/202610 Trash Wheel Collection Data.xlsx", sheet = 2, range = "A2:M139", col_names = TRUE, na = c("", ".", "NA")))|>
  janitor::clean_names()|>
  mutate(vessel = "prof", plastic_bottles = as.double(plastic_bottles)) |>
  relocate(vessel)

gwynnda_df = (read_excel("./data/202610 Trash Wheel Collection Data.xlsx", sheet = 4, range = "A2:L394", col_names = TRUE, na = c("", ".", "NA"))) |>
  janitor::clean_names()|>
  mutate(vessel = "gwyn")|>
  relocate(vessel)
```
## join datasets
```{r}
fulltrash_df = bind_rows(mrtrash_df, proftrash_df, gwynnda_df)
```

## Dataset information 
This data set includes `r nrow(fulltrash_df)` observations. It contains the following variables: `r colnames(fulltrash_df)`. These variables describe the trash collected by each trash vessel across the years. 

The total weight of trash collected by Professor Trash Wheel was `r fulltrash_df |> 
filter(vessel == "prof")|>
summarise(total_sum = sum(weight_tons, na.rm = TRUE))` tons. 

The total number of cigarette buts collected by Gwynnda in June of 2022 was `r fulltrash_df |> filter(vessel == "gwyn", year == "2022", month == "June") |> summarise( total_sum = as.integer(sum(cigarette_butts, na.rm = TRUE)))`. 

# Problem 3 

## import, clean & tidy baseline data
```{r}
library(tidyverse)
mcibase_df = read_csv("./data/MCI_baseline.csv", skip = 1, na=c("."))|>
  janitor::clean_names()|>
  rename (baseline_age = current_age)|>
  remove_rownames()|>
  mutate (
    sex = factor(sex, levels = c(1,0), labels = c("male", "female")),
    apoe4 = factor(apoe4, levels = c(1,0), labels = c("carrier", "noncarrier"))
          )|>
 filter(age_at_onset>baseline_age |is.na(age_at_onset))
```

## Dataset Information 

This dataset was imported and variable names were updated. Sex and Apoe4 variables were changed to factors and finally the dataset was filtered to remove any participants who had MCI diagnosis at baseline. Thus, the final dataset only includes those who met inlcusion criteria for the study. It includes four main variables: age at baseline, sex of participant, years of education, apoe4 status (carrier/noncarrier), and MCI age onset, with the following variable names: `r colnames(mcibase_df)` and has `r nrow(mcibase_df)` observations. 

There were `r nrow(mcibase_df)` participants recruited, and of them, `r mcibase_df|>
filter(!is.na(age_at_onset))|> nrow()` were diagnosed with MCI. 

The average age at baseline was `r mcibase_df |> 
summarise(round(mean(baseline_age, na.rm = TRUE), digits = 2))` years. 

The proportion of women in the study who are APOE4 carrriers is `r mcibase_df |> filter (sex == "female") |> summarise(round(mean(apoe4 == "carrier", na.rm = TRUE), digits = 2))`.


## add biomarker data

```{r}
mciamy_df = read_csv("./data/mci_amyloid.csv", skip = 1, na=c(".", "NA", "Na"))|>
  janitor::clean_names() |>
  remove_rownames() |>
  rename (id = study_id, amyratio_base = baseline, amyratio_time_2 = time_2, amyratio_time_4 = time_4, amyratio_time_6 = time_6, amyratio_time_8 = time_8)
```
The data was imported, specifying missing values to maintain numeric variables, and names were cleaned. The study id variable name was changed from study_id to id to be consistent across study datasets, and other variable names were changed to better reflect content. The data includes `r nrow(mciamy_df)` observations and, in additon to study id, 5 variables that report the amyloid beta 42/40  ratio across 5 timepoints for the following variables: `r colnames(mciamy_df)`. 

The amyloid dataset has `r nrow(mciamy_df)`observations while the baseline dataset only has `r nrow(mcibase_df)` observations, indicating there are participants that only appear in one dataset (the amyloid dataset).

## join datasets
```{r}
mcifull_df = left_join(mcibase_df, mciamy_df, by = "id")

write_csv (mcifull_df, "hw2_mci_joined.csv")
```

The combined dataset contains only participants who met inclusion criteria for the study; participants who had MCI at baseline were excluded. It includes baseline demographics of age at baseline, sex, years of education, apoe4 carrier status and measures of amyloid beta 42/40 ratios across five timepoints for the following variables `r colnames(mcifull_df)`, and has a total of `r nrow(mcifull_df)` participants (observations). 
