---
title: "p8015_hw2_cmh2275"
author: "cmh2275"
date: "2026-10-03"
output: github_document
---

# Problem 1 

## import data

```{r}
library(tidyverse)

options(tibble.print_min = 5)

transit_df = read_csv (file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
                       col_types = cols(Route8 ="c", Route9 = "c", Route10 = "c", Route11 = "c"), 
                       na=c("NA",".","")) |>
        janitor::clean_names() |>
  select(
    line, station_name, station_latitude, station_longitude, 
    starts_with("route"), entry, exit_only, vending, entrance_type, 
    ada) |> 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
  
```

## Paragraph re: data

This dataset includes information about the NYC transit station's entrances and exits across locations and routes. In additon to informaiton about the division, line, station names, station locations, and routes, it provides information about entrance types, entry, exit, vending, staffing and staff hours, ada compliance, if there is free crossover, cross streets, and entrance locations, with a total of 32 variables. The dataset was reduced to include 19 variables: line, station name, station latitude, station longitude, routes 1-11, vending, entrance type, entry, and ada compliance.The entry variable was converted from a character yes/no variable to a logical variable. There are 19 columns and 1868 rows 

## questions

```{r}
transit_df |> 
  select(station_name, line) |> 
  distinct()
```

*There are 465 distinct stations.

```{r}
transit_df |> 
  filter(ada == TRUE) |> 
  select(station_name, line) |> 
  distinct()
```

* 85 Stations are ADA compliant. 

```{r}
transit_df|> 
  filter(vending == "NO") |> 
  pull(entry) |> 
  mean()
```
* The proportion of stations entrances/exits without vending that allow entrance is 0.377. 
```{r}
transit_df |> 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") |> 
  filter(route == "A") |> 
  select(station_name, line) |> 
  distinct()

transit_df |> 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") |> 
  filter(route == "A", ada == TRUE) |> 
  select(station_name, line) |> 
  distinct()
```
* Sixty distinct stations serve the A train. 
* Seventeen of the sixty stations that serve the A train are ADA compliant.  

# Problem 2 

