---
title: "p8015_hw2_cmh2275"
author: "cmh2275"
date: "2026-10-03"
output: github_document
---

# Problem 1 

## import data

```{r}
library(tidyverse)

options(tibble.print_min = 5)

transit_df = read_csv (file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
                       na=c("NA",".",""))

```
## clean names
```{r}
transit_df = 
  janitor::clean_names(transit_df)

```

## select columns and change variable
```{r}
transit_df = select(transit_df, line, station_name, station_latitude, station_longitude, route1, route2, route3, route4, route5, route6, route7, route8, route9, route10, route11, entrance_type, entry, ada) 
  
transit_df = mutate (transit_df, 
    entry = case_match(
          entry, 
            "YES" ~ TRUE,
            "NO" ~ FALSE))

```

## Paragraph re: data

This dataset includes information about the NYC transit station's entrances and exits across locations and routes. In additon to informaiton about the division, line, station names, station locations, and routes, it provides information about entrance types, entry, exit, vending, staffing and staff hours, ada compliance, if there is free crossover, cross streets, and entrance locations, with a total of 32 variables. The dataset was reduced to include 18 variables: line, station name, station latitude, station longitude, routes 1-11, entrance type, entry, and ada compliance.The entry variable was converted from a character yes/no variable to a logical variable. There are 18 columns and 1868 rows 
